Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
add all relevant citations

Computer networks run many critical services, and every second of downtime is
costly at best and dangerous at worst.  Yet, keeping these systems
running 24/7 is an enormous
challenge~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
%
While hardware faults, backhoes, power failures and natural disasters all pose
problems, studies have shown that human error in network
configuration is one of the leading causes of
outages~\cite{juniper-study,xx}.

@@

Worse, operators can never be sure that their templates correctly implement the policy. %
Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which policies they implement and for which topologies---other than those of the configurations obtained from instantiating them.
%
Thus, operators validate templates by testing some concrete instantiations using
simulation or verification tools~\cite{x,y}. This testing is insufficient because large networks evolve continuously, as
devices and links are added or taken offline for maintenance. Templates that
work for the current network topology, may or may not work for future
topologies, and it is intractable to test all possible future topologies. Updating the templates, if found to not work for a future topology,
may cause trickle-down, network-wide configuration changes, which is highly
disruptive.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{x,y,z,configassure,propane}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, these systems only understand concrete topologies, with no understanding of roles.  Even if two devices play the same role,
there is no guarantee that the systems will generate (syntactically) similar
configurations; if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of a handful of
role configurations. These systems become even more brittle when the network
evolves. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology.  The result may be a completely different
set of configurations.

@@

Second, \sysname can scale substantially better than its nearest competitor \propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgeries-popl16}.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

      Fattree~\cite{fattree} & -- & \cmark & \cmark & \cmark & \xmark  \\ \hline
      Facebook~\cite{facebook-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10~\cite{f10-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      VL2~\cite{vl2-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      BCube~\cite{bcube} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
      Dcell~\cite{dcell} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly~\cite{butterfly} & $n$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      HyperX~\cite{hyperx} & $L$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \vspace{-.6em}
  \caption{Fault-tolerance analysis precision for all topologies with shortest path routing.}
  \label{fig:analysis-precision}
  \vspace{-1em}
\end{figure}

@@

We evaluate the expressiveness of \sysname's abstractions and the precision of the fault-tolerance analysis on a range of different network topologies found both in the network literature and in production networks. The analysis is precise if it finds a tight lower-bound on the fault-tolerance (\IE, there is actually a concrete network with that fault tolerance).
For each topology in Figure~\ref{fig:analysis-precision} we define a suitable abstraction. Since many topologies have tunable parameters (\EG, the recursion depth $k$ for the Dcell topology), we list all such parameters that are fixed for the abstraction. For all topologies (Figure~\ref{fig:analysis-precision}), we implement a shortest path routing policy. Additionally, for clos-style topologies, we implement a valley-free routing policy (Figure~\ref{fig:analysis-precision2}) for comparison. In each case, we record the precision of the analysis in terms of whether or not it is able to accurately prove reachability and/or the correct number of disjoint paths between some or all pairs of nodes from a fixed source.

@@

We also model several recursive topologies. These include the BCube~\cite{bcube} and Dcell~\cite{dcell} topologies. Each topology includes a recursion depth parameter ($k$), which we fix. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but not from increasing the recursion depth.
For BCube, we model each tier of the datacenter as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.
The analysis is able to accurately determine reachability (i.e., at least 1 disjoint path) for each recursive topology, but is overly conservative in determining k-disjoint paths. For the BCube topology, it is able to to determine the correct number of disjoint paths between some pairs of nodes, but not all.

@@

\para{Network synthesis} Many recent systems focus on allowing operators to specify their policies at a high level. These can be classified into two classes. The first class targets networks based on SDN (software-defined networking) and directly generates the forwarding rules that tell routers how to forward packets~\cite{x,y,z}. The second class, and to which our work belongs, targets conventional networks and generates router configuration, based on which the routers interact and compute forwarding rules~\cite{propane,narain:lisa05,narain+:configassure}.
Almost all large networks (old or new) use the conventional paradigm because that scales better~\cite{bgp-in-dc}.

@@

\para{Topology design} Network topology design, especially for data centers, is another active research area, with researchers exploring designs with different properties~\cite{fattree,facebook-fattree,f10-fattree,vl2,bcube,dcell,hyperx,butterfly}.  We do not develop new designs but develop abstractions that can capture these designs (and their evolution). %and algorithms that can analyze the fault tolerance of abstract topologies.

While not the focus of this paper, our tools can enable operators to pick the right design for their networks, more quickly and with stronger guarantees. Today, when picking the design, operators analyze its multiple concrete instantiations because available analysis tools only work with concrete topologies~\cite{condor}. This process is time-consuming and not fool-proof; analyzing large, concrete topologies (with potentially thousands of nodes) takes a lot of time and the analysis does not cover the entire space of topologies that may arise.

\para{Network verification}
A complementary approach to reducing configuration errors is by analyzing the forwarding rules or configuration of the network to ensure the absence of (a class of) bugs~\cite{anteater,hsa,feamster+:rcc,ipassure,batfish,bagpipe,arc,era,symmetries}. Our focus, in contrast, is on a correct-by-construction approach. It has the downside that operators must learn to express policies in a new language, but it saves them from the challenging and time-consuming task of configuration generation and provides stronger guarantees on correctness.

>>>>

Computer networks run many critical services, and every second of downtime is
costly at best and dangerous at worst.  Yet, keeping these systems
running 24/7 is an enormous
challenge~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
%
While hardware faults, backhoes, power failures and natural disasters all pose
problems, studies have shown that human error in network
configuration is one of the leading causes of
outages~\cite{juniper-study,yankee-study}.

@@

Worse, operators can never be sure that their templates correctly implement the policy. %
Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which policies they implement and for which topologies---other than those of the configurations obtained from instantiating them.
%
Thus, operators validate templates by testing some concrete instantiations using
simulation or verification tools~\cite{feamster+:rcc, batfish, era, arc, bagpipe}. This testing is insufficient because large networks evolve continuously as
devices and links are added or taken offline for maintenance. Templates that
work for the current topology, may or may not work for future
topologies, and it is intractable to test all possible future topologies. Updating the templates, if found to not work for a future topology,
may cause trickle-down, network-wide configuration changes, which is highly
disruptive.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{narain+:configassure,propane}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, these systems only understand concrete topologies, with no understanding of roles.  Even if two devices play the same role,
there is no guarantee that the systems will generate (syntactically) similar
configurations; if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of a handful of
role configurations. These systems become even more brittle when the network
evolves. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology.  The result may be a completely different
set of configurations.

@@

Second, \sysname can scale substantially better than its nearest competitor \propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

      Fattree~\cite{fattree} & -- & \cmark & \cmark & \cmark & \xmark  \\ \hline
      Facebook~\cite{facebook-fattree} & -- & \cmark & \cmark & \cmark & \cmark \\ \hline
      F10~\cite{f10-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      VL2~\cite{vl2-fattree} & -- & \cmark & \cmark & \cmark & \xmark \\ \hline
      BCube~\cite{bcube} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
      DCell~\cite{dcell} & $k$ & \cmark & \cmark & \xmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly~\cite{butterfly} & $n$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      HyperX~\cite{hyperx} & $L$ & \cmark & \cmark & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \vspace{-.6em}
  \caption{Fault-tolerance analysis precision for all topologies with shortest path routing.}
  \label{fig:analysis-precision}
  \vspace{-1em}
\end{figure}

@@

We evaluate the expressiveness of \sysname's abstractions and the precision of the fault-tolerance analysis on a range of different network topologies found both in the network literature and in production networks. The analysis is precise if it finds a tight lower-bound on the fault-tolerance (\IE, there is actually a concrete network with that fault tolerance).
For each topology in Figure~\ref{fig:analysis-precision} we define a suitable abstraction. Since many topologies have tunable parameters (\EG, the recursion depth $k$ for the DCell topology), we list all such parameters that are fixed for the abstraction. For all topologies (Figure~\ref{fig:analysis-precision}), we implement a shortest path routing policy. Additionally, for clos-style topologies, we implement a valley-free routing policy (Figure~\ref{fig:analysis-precision2}) for comparison. In each case, we record the precision of the analysis in terms of whether or not it is able to accurately prove reachability and/or the correct number of disjoint paths between some or all pairs of nodes from a fixed source.

@@

We also model several recursive topologies. These include the BCube~\cite{bcube} and DCell~\cite{dcell} topologies. Each topology includes a recursion depth parameter ($k$), which we fix. For a recursive topology with depth $k$, we model it as an abstract topology consisting of a pod to represent all depth $k-1$ subcomponents. This allows for safe expansion within a subcomponent, but not from increasing the recursion depth.
For BCube, we model each tier of the datacenter as a separate role. Figure~\ref{fig:example-abstractions} shows an example of a BCube abstraction for $k=1$.
The analysis is able to accurately determine reachability (i.e., at least 1 disjoint path) for each recursive topology, but is overly conservative in determining k-disjoint paths. For the BCube topology, it is able to to determine the correct number of disjoint paths between some pairs of nodes, but not all.

@@

\para{Network synthesis} Many recent systems focus on allowing operators to specify their policies at a high level. These can be classified into two classes. The first class targets networks based on SDN (software-defined networking) and directly generates the forwarding rules that tell routers how to forward packets~\cite{foster:merlin,fattire,netgen}. The second class, and to which our work belongs, targets conventional networks and generates router configuration, based on which the routers interact and compute forwarding rules~\cite{propane,narain:lisa05,narain+:configassure}. 
Almost all large networks (old or new) use the conventional paradigm because that scales better~\cite{bgp-in-dc}.

@@

\para{Topology design} Network topology design, especially for data centers, is another active research area, with researchers exploring designs with different properties~\cite{fattree,facebook-fattree,f10-fattree,vl2-fattree,bcube,dcell,hyperx,butterfly}.  We do not develop new designs but develop abstractions that can capture these designs (and their evolution).

While not the focus of this paper, \sysname can enable operators to pick the right design for their networks, more quickly and with stronger guarantees. Today, when picking a design, operators analyze its multiple concrete instantiations because available analysis tools work with concrete topologies~\cite{condor}. This process is time-consuming and not fool-proof; analyzing large, concrete topologies (with potentially thousands of nodes) takes a lot of time and the analysis does not cover the entire space of topologies that may arise.

\para{Network verification}
A complementary approach to reducing configuration errors is by analyzing the forwarding rules or configuration of the network to ensure the absence of (a class of) bugs~\cite{anteater,hsa,veriflow,feamster+:rcc,ipassure,batfish,bagpipe,arc,era,surgery}. Our focus, in contrast, is on a correct-by-construction approach. It has the downside that operators must learn to express policies in a new language, but it saves them from the challenging and time-consuming task of manual configuration generation and correct-by-construction guarantees.
